---
title: "Aggregate landings to another footprint"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Aggregate landings to another footprint}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

```

```{r setup}
library(comlandr)
```

Specifying the arguments `aggArea = T` (in `get_comland_data()`) and supplying a data frame, `userAreas`, the footprint of the returned landings can be changed by aggregating the landings over [statistical areas](https://www.fisheries.noaa.gov/resource/map/greater-atlantic-region-statistical-areas) (the units of reporting). 

* Any statistical areas not defined in `userAreas` are assigned to an `OTHER` category
* Landings and value are aggregated to areas defined in `userAreas`
* To aggregate NAFO data is more complicated since the reporting footprint differs from that of STOCKEFF
  * A proportion of the landings in each NAFO division is assigned to `userAreas` based on the overlap of statistical areas in `userAreas` with the NAFO division

For a list of Statistical Areas see `get_areas()`
