---
title: "comlandr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{comlandr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(comlandr)
```

## Connect to the database

Once you are granted permission to access the database and have Oracle's instant Client installed you can make a connection as follows:

```channel <- dbutils::connect_to_database(server="servername",uid="yourUsername")```

`channel` is an object inherited from the DBIConnection-class. This object is passed as an argument to functions in `comnlandr`

## Pull Raw Data

Pulling the raw data takes up to 15 minutes to complete. see `get_comland_data` for details. To pull the commercial data with shellfish represented as meat weight (landed = "y")

```
data <- get_comland_data(channel,landed="y",endyear=2019,out.dir=here::here())
```

`data` is a data frame (see `?get_comland_data`) for more details

This data is considered the raw data and does not contain any corrections that stock assessment scientists may do. For example,  market categories representing parts of a fish are not removed.

## Corrected Raw Data

The raw data is processed to in several ways

* Fish parts are removed
* Separate little and winter skate aggregate based on survey proportions
* Separate silver hake and offshore hake based on survey proportions
* Herring data are pulled from Maine database
* catch with missing fields (AREA, Size, Gear) are assigned based on trips with similar characteristics (Palmer ????)
* Additional Foreign landings are added from NAFO*

To calculate the processed Raw data with default arguments:

``` comland(channel) ```


