---
title: "comlandr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{comlandr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(comlandr)
```

## Prequisite

You will need to have `ROracle` installed and configured on your machine

## Connect to the database

You can make a connection as follows:

```channel <- dbutils::connect_to_database(server="servername",uid="yourUsername")```

`channel` is an object inherited from the `ROracle` package. This object is passed as an argument to functions in `comlandr`

## Pull Raw Data

Pulling the raw data takes up to 15 minutes to complete. see [`get_comland_raw_data`](../reference/get_comland_raw_data.html) for details. 

To pull the raw (unprocessed) commercial data for:

* Years 1964 through 2023
* With shellfish represented as landed weight (as opposed to live weight)
* By Statistical Area

```
data <- get_comland_raw_data(channel,filterByYear = 1964:2023)
```

The resulting `data` object is a list containing several fields. See [`get_comland_raw_data`](../reference/get_comland_raw_data.html) for details. 

The data is considered the raw data and does not contain any corrections that stock assessment scientists may do. For example, some market categories representing parts of a fish are not removed.

Note: Herring and Menhaden data are underrepresented in this raw data pull. Herring data is pulled separately from another database. Menhaden are excluded due to confidentiality concerns

## Processed Raw Data

To pull the processed raw data, use [`get_comland_data`](../reference/get_comland_data.html)

```
data <- get_comland_data(channel,filterByYear = 1964:2023)
```

This will pull the raw data (above) then process it depending on the input arguments supplied.

The options in the processing are:

* Include Herring data
* Remove fish parts
* Include foreign landings from NAFO
* Adjust species value based on a reference year
* Separate skate and hake landings by species based on survey proportions
* Aggregate landings by user defined groupings of Statistical areas
* Aggregate gear codes based on user supplied fleets
* Attempt to assign values to missing data

See the articles section for in depth explanation of the components of `comlandr`
