---
title: "Landings by foreign vessels in US waters"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{landingsNafo}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(comlandr)
```

Landings in US waters by foreign vessels are reported to NAFO. This data is obtained directly from NAFOs website using `get_foreign_data()`. 

* Details regarding the types of data available see `vignette("ForeignCatch")`.
* For details regarding the reporting areas, the countries reporting and the species caught see `vignette("ForeignCatchCompare21A21B")`

Data from NAFO require considerable processing and several assumptions are made before merging with the US landings data from STOCKEFF. 

* The spatial domain for reporting does not fully overlap with the statistical areas in which US data is reported.
* Species codes and names differ between the two data sets

## Processing steps

* US data is removed
* Data with missing `MONTH` field are assigned, `MONTH` = 0 and `QY`= 1 
* Data is aggregated over country codes
* Data is filtered to keep only Division codes = c(47, 51:56, 61:63) (See `vignette("ForeignCatchCompare21A21B")`)
* Missing data from 21B data is substituted with 21A and assigned `MONTH` = 0 and `QY`= 1 
* `cfdbs.CFSPP` lookup table is used to match NAFO species codes with `NESPP3` codes
  * Codes present in the NAFO data but not defined in lookup tables are matched manually
  * NAFO species codes used as catch alls, for example Groundfish (Not specified), are assigned to corresponding `NESPP3` codes
* All shellfish are reported as live weight. A conversion to meat weight is applied if `useLanded` is supplied.
* Herring data is removed (since the State of Maine database account for ALL Herring landings, see `vignette("landingsHerring")`)
* `cfdbs.GEAR` lookup table is used to match NAFO gears with `NEGEAR` codes
  * Mismatched codes are then manually adjusted
* Vessel tonnage class codes in NAFO data are matched to `TONCL2` codes.
* Additional missing fields (`SPPVALUE`,`UTILCD`, `MESHCAT`,`MARKET_CODE`) are set to missing
