---
title: "Pull US landings data from STOCKEFF"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Pull US landings data from STOCKEFF}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(comlandr)
```

US landings by trip are pulled from STOCKEFF via `get_comland_raw_data()` and grouped by year, month, negear, toncl2, nespp3, nespp4, area, utilcd, mesh, market_code from which spplivlb, sppvalue, and spplndlb are summed.

* Mesh category (`MESHCAT`) is then defined as either `SM` (<= 3 inches) or `LG` (> 3 inches) depending on whether a trip reported a mesh size

* Depending on the value of the argument `useLanded`, meat weight or live weight is used for shellfish species (NESPP3 codes for shellfish are 743:800). Use `get_species()` or `get_species_itis()` to associate common or scientific names to these shellfish codes

* Fish parts are removed. For a list of removed fish part codes, see `vignette("fishPartsRemoved")`

* All weight is converted from lbs to metric tons
