---
title: "missingData"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{missingData}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(comlandr)
```

The commercial data isn't perfect. There are trips where the AREA in which the catch was taken is missing. There are trips in which the gear type used (NEGEAR) is missing, and there are trips in which the MONTH of the trip is missing. It is important to know these, since ALL landings need to be accounted for, especially for Stock assessments. An R implementation of the Palmer (2008) algorithm is applied to the raw data pull to assign values to this missing data based on "similar" trips.

* Missing fields (MONTH, NEGEAR, AREA) are imputed using HY, QY, MONTH, NEGEAR, TONCL2, AREA from trips with known values

* Future releases may integrate STOCKEFFs implementation of this algorithm to further align `comlandr` with products used in stock assessments
