---
title: "Notes about comlandR.r"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Notes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(comland)
```

For details of its use see the [tech-doc](https://noaa-edab.github.io/tech-doc/comdat.html)

1. Either pulls data from data tables (~15 mins) or reads in from a previous pull
1. aggregate landings by stat area or EPU (multiple stat areas)
1. include foreign landings as defined by NAFO
1. adjusts value of species for inflation

## output

``` {r eval = T, echo = T}
comland <- readRDS(here::here("output","comland_meatwt_deflated_EPU.Rds"))
head(comland)
```

Column Headings:

* QY - Quarter year (1,2,3,4)
* GEAR - aggreated gear type comprised of NEGEAR codes (See below)

``` {r gears, eval = T, echo = T}
GEARs
```

* SIZE - 
* EPU - Ecological Production Unit comprised of AREA codes (See below)
``` {r epus, eval = T, echo = T}
EPUs
```

* UTILCD - 
* SPPLIVMT - live weight of species
* SPPVALUE - value of species (?? how calculated)
* US - US landings or foreign


