---
title: "Calculating present day value"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{speciesValue}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(comlandr)
```

The `SPPVALUE` field in the STOCKEFF database is the value of the catch on the date landed. To compare `SPPVALUE` over time, these values need to be converted to present day dollar amounts. This is achieved using the [Implicit Price Deflator](https://fred.stlouisfed.org/series/GDPDEF) of Gross Domestic Product (GDPDEF) from the Federal Reserve Bank of St Louis. This index is calculated on a quarterly basis. In `comlandr` we hold the deflator constant for each quarter.

* `SPPVALUE` in STOCKEFF is only available from 1981 to present day. 
* `comlandr` allows the user to adjust the value to any reference date required. See arguments `refMonth`, `refYear` in `get_comland_data()`

